Bow shocks, bow waves, and dust waves. IV. Shell shape statistics

### 2019-05-02 papers/Papers1234/sec-obs-intro.tex
Make a start on the discussion section of Paper IV

The shapes of stellar bow shocks are frequently compared with what has
become known as the \textit{wilkinoid} surface \citep{Cox:2012a},
which is the result of the idealized interaction between a spherical
wind and a plane-parallel stream in the hydrodynamic thin-shell
approximation.  Numerical approximations to this shape were used by
various authors \citep{Baranov:1971a, Mac-Low:1991a} before an elegant
analytic solution was found by \citet{Wilkin:1996a} and extended to
the case of interaction between two spherical winds
\citep{Canto:1996}.  In \citet[hereafter, Paper~0]{Tarango-Yong:2018a}
the study of bow shock shapes and their projection on the plane of the
sky was formalized. The term \textit{cantoid} was introduced for the
\citet{Canto:1996} family of shapes, together with \textit{ancantoid}
for a generalization to the case where one of the winds is
anisotropic, as is appropriate for the proplyds.  In addition, Paper~0
proposed the use of two dimensionless parameters, \textit{planitude}
and \textit{alatude}, to describe a general bow shock shape.  The
planitude, \(\Pi = R_c / R_0\), measures the flatness of the bow shock
apex (see Fig.~1 of Paper~0 for explanation of apex and other
descriptive terms), where \(R_0\) is the star--apex distance and
\(R_c\) is the radius of curvature measured at the apex.  The alatude,
\(\Lambda = R_{90}/R_0\), measures the openness of the bow shock wings,
where \(R_{90}\) is the lateral size of the bow, measured from the
star in the direction perpendicular to the star--apex direction.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/obs-shape-terminology}
  \caption[Terminology]{Terminology employed in this paper to describe
    bow shock shapes, following \citet{Tarango-Yong:2018a}.}
  \label{fig:obs-shape-terminology}
\end{figure}

The shapes of stellar bow shocks are frequently compared with what has
become known as the \textit{wilkinoid} surface \citep{Cox:2012a},
which is the result of the idealized interaction between a spherical
wind and a plane-parallel stream in the hydrodynamic thin-shell
approximation.  Numerical approximations to this shape were used by
various authors \citep{Baranov:1971a, Mac-Low:1991a} before an elegant
analytic solution was found by \citet{Wilkin:1996a} and extended to
the case of interaction between two spherical winds
\citep{Canto:1996}.  In \citet[hereafter, Paper~0]{Tarango-Yong:2018a}
the study of bow shock shapes and their projection on the plane of the
sky was formalized. The term \textit{cantoid} was introduced for the
\citet{Canto:1996} family of shapes, together with \textit{ancantoid}
for a generalization to the case where one of the winds is
anisotropic, as is appropriate for the proplyds.  In addition, Paper~0
proposed the use of two dimensionless parameters, \textit{planitude}
and \textit{alatude}, to describe a general bow shock shape, which we
illustrate in Figure~\ref{fig:obs-shape-terminology}.  The planitude,
\(\Pi = R_c / R_0\), measures the flatness of the bow shock apex, where
\(R_0\) is the star--apex distance and \(R_c\) is the radius of
curvature measured at the apex.  The alatude,
\(\Lambda = R_{90}/R_0\), measures the openness of the bow shock wings,
where \(R_{90}\) is the lateral size of the bow, measured from the
star in the direction perpendicular to the star--apex direction.

### 2019-05-02 papers/Papers1234/sec-quadrics-observations.tex
Switch to Kuiper test instead of K-S
Make a start on the discussion section of Paper IV

The horizontal axis corresponds to the shape of the head of the bow
shock near its apex, ranging from sharper, pointier shapes with
\(\Pi < 1\) to flatter, snubber shapes with \(\Pi \gg 1\), where it must be
understood that all judgments of sharpness/flatness are with respect
to the axial separation, \(R_0\), between the source and the bow shock
apex.  The vertical axis corresponds to the shape of the bow shock
wings, ranging from closed ``C'' shapes for smaller values of
\(\Lambda\) to open ``V'' shapes for larger values of \(\Lambda\).  The background
shading shows the regions of the \(\Pi\)--\(\Lambda\) plane occupied by simple
quadric shapes (see \S~4 of Paper~0): oblate spheroids (light gray),
prolate spheroids (dark gray), and hyperboloids (white). The boundary
between closed and open corresponds to the paraboloids, and is shown
by the solid curved line that divides the dark gray and white regions
of the graph.

@@

We obtain a third sample of bow shocks from a catalog of stationary
emission line arcs in the Orion Nebula \citep{Bally:2000a}, which are
detected via their H\(\alpha\) emission in HST~ACS surveys of the nebula
\citep{Bally:2006a, Robberto:2013a}.  Out of a total of 73 such
objects \citep{Gutierrez-Soto:2015a} we have selected 18 where the
observations are of sufficient quality to reliably measure the
planitude and alatude.  Images in the ACS f658n filter of these
sources are shown in Figure~\ref{fig:ll-arcs}.  Some of the sources
were identified as LL~Orionis-type objects by \citet{Bally:2001a} and
are identified as LL~1--6. The remaining sources are labeled by their
coordinate-based designation according to the system of
\citet{ODell:1994a}.  The determination of the bow shock shape for
each source was carried out as described in \S~7 of Paper~0, using the
``ridge'' method for manually tracing the arc (see Fig.~28 of
Paper~0).  The planitude and alatude are found by circle fitting using
the algorithm described in Paper~0's Appendix~E, as implemented in the
python program
\texttt{circle-fit.py}.\footnote{\url{https://github.com/div-B-equals-0/circle-fit}}
The only adjustable parameter of the algorithm is \(\Delta\theta\), the maximum
angle from the bow shock axis of points that are included in the
circle fit.  We tested all values of \(\Delta\theta\) in \ang{5} increments from
\ang{45} to \ang{80} and found that \(\Delta\theta = \ang{55}\) to \ang{70}
gives stable results for all sources.  The accuracy of the fitted
planitudes can be estimated from the dispersion in values of \(\Pi\) for
the same source with different \(\Delta\theta\), which we find to be
\(\approx 10\%\).  The fitted alatudes have a much smaller dispersion
\(\approx 1 \%\), but in this case the accuracy is limited by the asymmetry
of the wings, as characterized by \(\Delta\Lambda/\Lambda\) with an RMS value of
\(\approx 15\%\).

>>>>

The horizontal axis corresponds to the shape of the head of the bow
shock near its apex, ranging from sharper, pointier shapes with
\(\Pi < 1\) to flatter, snubber shapes with \(\Pi \gg 1\), where it must be
understood that all judgments of sharpness/flatness are with respect
to the axial separation, \(R_0\), between the source and the bow shock
apex (see Fig.~\ref{fig:obs-shape-terminology}).  The vertical axis
corresponds to the shape of the bow shock wings, ranging from closed
``C'' shapes for smaller values of \(\Lambda\) to open ``V'' shapes for
larger values of \(\Lambda\).  The background shading shows the regions of
the \(\Pi\)--\(\Lambda\) plane occupied by simple quadric shapes (see \S~4 of
Paper~0): oblate spheroids (light gray), prolate spheroids (dark
gray), and hyperboloids (white). The boundary between closed and open
corresponds to the paraboloids, and is shown by the solid curved line
that divides the dark gray and white regions of the graph.

@@

We obtain a third sample of bow shocks from a catalog of stationary
emission line arcs in the Orion Nebula \citep{Bally:2000a}, which are
detected via their H\(\alpha\) emission in HST~ACS surveys of the nebula
\citep{Bally:2006a, Robberto:2013a}.  Out of a total of 73 such
objects \citep{Gutierrez-Soto:2015a} we have selected 18 where the
observations are of sufficient quality to reliably measure the
planitude and alatude.  Images in the ACS f658n filter of these
sources are shown in Figure~\ref{fig:ll-arcs}.  Some of the sources
were identified as LL~Orionis-type objects by \citet{Bally:2001a} and
are identified as LL~1--6. The remaining sources are labeled by their
coordinate-based designation according to the system of
\citet{ODell:1994a}.  The determination of the bow shock shape for
each source was carried out as described in \S~7 of Paper~0, using the
``ridge'' method for manually tracing the arc (see Fig.~28 of
Paper~0).  In some sources, such as LL~1, LL~6, 109--246, and 4468--605,
a high-velocity jet flow is also seen to issue from the star
\citep{Bally:2006a, Henney:2013a}.  We took care to when tracing the
arcs of such sources to avoid regions where jet knot emission is
projected on the bow shock shell.

The planitude and alatude are found by circle fitting using the
algorithm described in Paper~0's Appendix~E, as implemented in the
python program
\texttt{circle-fit.py}.\footnote{\url{https://github.com/div-B-equals-0/circle-fit}}
The only adjustable parameter of the algorithm is \(\Delta\theta\), the maximum
angle from the bow shock axis of points that are included in the
circle fit.  We tested all values of \(\Delta\theta\) in \ang{5} increments from
\ang{45} to \ang{80} and found that \(\Delta\theta = \ang{55}\) to \ang{70}
gives stable results for all sources.  The accuracy of the fitted
planitudes can be estimated from the dispersion in values of \(\Pi\) for
the same source with different \(\Delta\theta\), which we find to be
\(\approx 10\%\).  The fitted alatudes have a much smaller dispersion
\(\approx 1 \%\), but in this case the accuracy is limited by the asymmetry
of the wings, as characterized by \(\Delta\Lambda/\Lambda\) with an RMS value of
\(\approx 15\%\).

### 2019-05-02 papers/Papers1234/sec-standing-wave.tex
Make a start on the discussion section of Paper IV

>>>>

In this section, we discuss the physical implications of our empirical
findings regarding bow shock shapes.  Our most reliable result is the
average shape of the OB bow shocks from the 227 MIPSGAL sources with
quality rating of 3~stars or higher.  This yields mean values of
\(\Pi = 1.78 \pm 0.06\) and \(\Lambda = 1.72 \pm 0.02\), or median values of
\(\Pi = 1.57\) and \(\Lambda = 1.69\).  The uncertainty quoted on the mean
values is the ``standard error of the mean'':
\(\text{sem} = \sigma / \sqrt{n}\), where \(\sigma\) is the rms dispersion and
\(n\) is the number of sources.  Note that in the case of the
planitude \(\text{sem}(\Pi) = 0.06\) is considerably smaller than
\(\text{mean}(\Pi) - \text{median}(\Pi) = 0.21\), so the latter would be a
more conservative estimate of the uncertainty.\footnote{This is
  because the distribution of \(\Pi\) is approximately log-normal, which
  yields a significant tail towards high values when converted to
  linear space.}  These values can be compared with the predictions of
the thin-shell Wilkinoid model \citep{Wilkin:1996a}, which are
\(\Pi = 1.67\), \(\Lambda = 1.73\) when the bow shock axis lies in the plane
of the sky (following Paper~0, this is defined as the zero point of
the inclination angle, \(i\)).  When the axis is inclined, both
planitude and alatude are predicted to decrease but not by very much,
tending to \(\Pi = 1.5\), \(\Lambda = 1.63\) as
\(\abs{i} \to \ang{90}\) (see \S~5.3 of Paper~0).  The median observed
value\footnote{In the presence of outliers, the median is a more
  robust estimate of the central tendency than is the mean.} falls
squarely inside this range for both the planitude and alatude, which
is a remarkable triumph for the \citet{Wilkin:1996a} model.

On the other hand, turning now to the \emph{variety} of bow shock
shapes, we see that the wilkinoid can no longer explain our results.
The rms dispersions of the planitude and alatude distributions for the
MIPSGAL sources are \(\sigma(\Pi) = 0.87\) and
\(\sigma(\Lambda) = 0.30\) (Tab.~\ref{tab:big-p}), which are respectively 5 times
and 3 times larger than the total range of variation of \(\Pi\) and
\(\Lambda\) predicted for the wilkinoid surface.  Although some of the
dispersion is due to uncertainties in the observations and the fitting
algorithm, this contribution is expected to be small, especially for
the larger, well-resolved sources, for which systematic uncertainties
in the methods for determining \(\Pi\) and \(\Lambda\) will
dominate. Conservative upper limits to the relative size of these
uncertainties were estimated in \S~7 of Paper~0 to be \(< 20\%\) for
\(\Pi\) and \(< 10\%\) for \(\Lambda\), whereas the observed dispersions are
roughly twice as large: \(\sigma(\Pi)/\Pi = 55\%\) and \(\sigma(\Lambda)/\Lambda = 18\%\).
